\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{makecell}
\usepackage{tabularx}

\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,tkz-euclide}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{quotes,angles,calc,intersections}

\usepackage{titling}
\usepackage{pdfpages}
\usepackage{color}
\usepackage{hyperref}

\usepackage{common}
\usepackage{linear}

\begin{document}

\title{Vector Calculus Notes}
\author{Brendan Burkhart}
\maketitle

\tableofcontents
\newpage

\section{Orthogonal Projection}

\begin{thm}\label{law-of-cosines}
    Let $\theta$ be the angle between non-zero vectors $A$ and $B$, and let $C = A - B$. The \emph{Law of Cosines} states that $\norm{C}^2 = \norm{A}^2 + \norm{B}^2 - 2\norm{A}\norm{B}\cos(\theta)$.
\end{thm}

\begin{figure}[ht!]
    \centering
    \begin{tikzpicture}[scale=1.4]
        \coordinate (O) at (0, 0);
        \coordinate (A) at (3, 5);
        \coordinate (B) at (6, 2);
        \coordinate (A') at ($(O)!(A)!(B)$);

        \draw [thick, ->] (O) -- (A);
        \draw [thick, ->] (O) -- (B);
        \draw [thick, dashed] (A') -- (A);
        \draw [very thick, red, ->] (O) -- (A');

        \draw [decorate,decoration={brace,amplitude=6pt,raise=3pt]}]
            (O) -- (A) node [black,midway,xshift=-0.6cm,yshift=0.3cm]
            {\footnotesize $A$};

        \draw [decorate,decoration={brace,amplitude=6pt,raise=3pt]}]
            (B) -- (O) node [black,midway,xshift=0.2cm,yshift=-0.5cm]
            {\footnotesize $B$};

        \draw [decorate,decoration={brace,amplitude=6pt,raise=3pt]}]
            (A) -- (A') node [black,midway,xshift=1.0cm, yshift=0.1cm]
            {\footnotesize $A - A'$};

        \draw [decorate,decoration={brace,amplitude=6pt,raise=3pt]}]
            (O) -- (A') node [black,midway,xshift=-0.2cm, yshift=0.5cm]
            {\footnotesize $A'$};

        \tkzMarkAngle[size=1cm,mark=|](B,O,A)
        \tkzLabelAngle[pos=1.2](B,O,A) {$\theta$}

        \tkzMarkRightAngle[size=0.4](A,A',O)
    \end{tikzpicture}
\caption{Orthogonal projection of $\vec{A}$ onto $\vec{B}$}
\label{fig:orthogonal-projection}
\end{figure}

Figure \ref{fig:orthogonal-projection} shows $A'$, the orthogonal projection of vector $A$ onto $B$. $\norm{A'}$ is the shortest distance between the head of $A$ and $B$.

\begin{thm}
    Let $\theta$ be the angle between non-zero vectors $A$ and $B$. Then $\cos\theta = \frac{A \cdot B}{\norm{A}\norm{B}}$.
\end{thm}

\begin{proof}\label{vector-angle}
    Let $C = A - B$. Then $C \cdot C = (A - B) \cdot (A - B)$, which can we expand to get $A \cdot (A - B) - B \cdot (A - B) = A \cdot A + B \cdot B - 2(A \cdot B)$. Therefore, $\norm{C}^2 = \norm{A}^2 + \norm{B}^2 - 2(A \cdot B)$. By the Law of Cosines \ref{law-of-cosines} and transitivity, it follows that $A \cdot B = \norm{A}\norm{B}\cos\theta$, so $\cos\theta = \frac{A \cdot B}{\norm{A}\norm{B}}$.
\end{proof}

\begin{exmp}
    Let $A = \left(0, 1\right)$, and $B = \left(2, 2\right)$. Then $\cos\theta = \frac{A \cdot B}{\norm{A}\norm{B}}$. Since $A \cdot B = 0 \cdot 2 + 1 \cdot 2 = 2$, $\norm{A} = \sqrt{0^2 + 1^2} = 1$, and $\norm{B} = \sqrt{2^2 + 2^2} = \sqrt{8} = 2\sqrt{2}$, we have $\cos\theta = \frac{2}{1 \cdot 2\sqrt{2}}$. This is equal to $\frac{\sqrt{2}}{2}$, so $\theta = \frac{\pi}{4}$.
\end{exmp}

\begin{exmp}
    Let $A = {1, 1}$ and $B = {1, -1}$. Then since $A \cdot B = 1 - 1 = 0$, we know $\cos\theta = 0$, which implies that $\theta = \frac{\pi}{2}$.
\end{exmp}

\begin{cor}\label{perpendicular-vectors}
    Non-zero $A$ and $B$ are perpendicular if and only if $A \cdot B = 0$.
\end{cor}

\begin{proof}
    Let $A$ and $B$ be non-zero perpendicular vectors. Then the angle ($\theta$) between $A$ and $B$ must be $\frac{\pi}{2}$ since they are perpendicular. By Theorem \ref{vector-angle}, $\cos\theta = \frac{A \cdot B}{\norm{A}\norm{B}}$. Since $A$ and $B$ are non-zero, $\norm{A}\norm{B}$ is non-zero, and since $\theta = \frac{\pi}{2}$, we know $\cos\theta = 0$. Therefore, $A \cdot B$ must equal zero.
\end{proof}

\begin{thm}\label{vector-projection}
    Let $A$ and $B$ be vectors, and $A'$ be the orthogonal projection of $A$ onto $B$. Then $A' = \frac{A \cdot B}{B \cdot B}B$.
\end{thm}

\begin{proof}
    Let $d \in \R$ such that $C = dB$. Then since $(A - dB) \cdot B = 0$, we know $A \cdot B - dB \cdot B = 0$, so $A \cdot B = d(B \cdot B)$. This implies that $d = \frac{A \cdot B}{B \cdot B}$. Since $C = dB$, we then have $C = \frac{A \cdot B}{B \cdot B}B$.
\end{proof}

\begin{exmp}
   Let $A = \left(2, 2\right)$, and $B = \left(1, 0\right)$. Then since $A \cdot B = 2$, and $B \cdot B = 1$, we know $A' = \frac{2}{1}B = \left(2, 0\right)$.
\end{exmp}

\section{Cross Product}

\begin{defn}
    The \emph{cross product} or \emph{vector product} of three-dimensional vectors $A = a_1i + a_2j + a_3k$ and $B = b_1i + b_2j + b_3k$ is written $A \times B$. \[A \times B = \begin{vmatrix}
        i & j & k \\ a_1 & a_2 & a_3 \\ b_1 & b_2 & b_3
    \end{vmatrix}\] The value of this determinant can be found via the Laplace formula (also called cofactor expansion) along any column or row. Expansion along the first row gives: \[A \times B =
    \begin{vmatrix}
        i & j & k \\ a_1 & a_2 & a_3 \\ b_1 & b_2 & b_3
    \end{vmatrix} =
    \begin{vmatrix}
        a_2 & a_3 \\ b_2 & b_3
    \end{vmatrix}i -
    \begin{vmatrix}
        a_1 & a_3 \\ b_1 & b_3
    \end{vmatrix}j +
    \begin{vmatrix}
        a_1 & a_2 \\ b_1 & b_2
    \end{vmatrix}k\] \[= (a_2b_3 - b_2a_3)i - (a_1b_3 - b_1a_3)j + (a_1b_2 - b_1a_2)k.\]
\end{defn}

\begin{rmk}
    The cross product is a non-commutative binary operation.
\end{rmk}

\begin{exmp}
    Let $A = 3i + j - k$, and $B = -6i - 2j - 2k$. Then \[A \times B =
    \begin{vmatrix}
        i & j & k \\ 3 & 1 & -1 \\ -6 & -2 & -2
    \end{vmatrix} = (-2 - 2)i - (-6 - 6)j + (-6 + 6)k = -4i + 12j + 0k.\]
\end{exmp}

\begin{thm}
    Let $A = a_1i + a_2j + a_3k$ and $B = b_1i + b_2j + b_3k$ be vectors. Then $A \times B$ is perpendicular to both $A$ and $B$.
\end{thm}

\begin{proof}
    Since $A \times B = (a_2b_3 - b_2a_3)i - (a_1b_3 - b_1a_3)j + (a_1b_2 - b_1a_2)k$, we know that $A \cdot (A \times B) = (a_1a_2b_3 - a_1b_2a_3) - (a_1a_2b_3 - b_1a_2a_3) + (a_1b_2a_3 - b_1a_2a_3)$. This is equal to $(a_1a_2b_3 - a_1a_2b_3) - (b_1a_2a_3 - b_1a_2a_3) + (a_1b_2a_3 - a_1b_2a_3)$, which is equal to $0$, so by Corollary \ref{perpendicular-vectors} we know $A$ is perpendicular to $A \times B$. Similarly, $B$ is perpendicular to $A \times B$.
\end{proof}

\begin{thm}
    Let $A$, $B$, and $C$ be vectors, and $x$ a real number. Then:
    \begin{itemize}
        \item $A \times B = 0$ if and only if $A = \vec{0}$ or $B = \vec{0})$, or $A$ is parallel to $B$.
        \item $A \times B = -(B \times A)$.
        \item $A \times (B + C) = (A \times B) + (A \times C)$.
        \item $(A + B) \times C = (A \times C) + (B \times C)$.
        \item $(\alpha A) \times B = A \times (\alpha B) = \alpha(A \times B)$.
    \end{itemize}
\end{thm}

\begin{proof}

\end{proof}

\begin{rmk}
    In a \emph{right-handed} coordinate system, the cross product will follow the \emph{right-hand rule}: on the right hand, index finger represents $A$, middle finger represents $B$, and then thumb represents $A \times B$. However, in a \emph{left-handed} coordinate system, the cross product will instead follow the similar \emph{left-hand rule}. The chirality of the cross product follows the chirality of the chosen basis, and is not inherently due to the definition of the cross product.
\end{rmk}

\begin{thm}
    Let $A$ and $B$ be two vectors, and $\theta$ be the angle between them. Then $\norm{A \times B} = \norm{A}\norm{B}\sin\theta$, which is the area of a parallelogram with side lengths $\norm{A}$ and $\norm{B}$ and angle $\theta$.
\end{thm}

\begin{thm}
    Let $(x_1, y_1, z_1)$ be a point, and $Ax + By + Cz + D = 0$ be a plane. Then the distance from the point to the plane is: \[\frac{\abs{Ax_1 + By_1 + Cz_1 + D}}{\sqrt{A^2 + B^2 + C^2}}.\]
\end{thm}

\begin{proof}
    The line with direction $(A, B, C)$ that passes through the point is normal to the plane and thus contains the shortest path from the point to the plane. Therefore, $A(x_1 + At) + B(y_1 + Bt) + C(z_1 + Ct) + D = 0$ gives the closest point on the plane to the point. $Ax_1 + By_1 + Cz_1 + D = -t(A^2 + B^2 + C^2)$. The shortest distance is thus \[\abs{\frac{t}{\norm{(A, B, C)}}} = \frac{\abs{Ax_1 + By_1 + Cz_1 + D}}{\sqrt{A^2 + B^2 + C^2}}.\]
\end{proof}

\section{Alternative Coordinate Systems}

There are many useful alternatives to the Cartesian coordinate system, such as the polar, cylindrical, and spherical coordinate systems.

In polar coordinates, two-dimensional points are represented by a radius $r$ and an angle $\theta$. The $r$ is the radius from the origin, and $\theta$ is the angle measured anticlockwise around the origin, starting from the positive half of the $x$-axis.

\begin{thm}
    Let $(x, y)$ be a point in Cartesian coordinates. Then $(r, \theta)$, where
    \begin{itemize}
        \item $r = \sqrt{x^2 + y^2}$
        \item $\theta = \arctan\frac{y}{x}$
    \end{itemize} gives the same point in polar coordinates.
\end{thm}

\begin{thm}
    Let $(r, \theta)$ be a point in polar coordinates. Then $(x, y)$, where
    \begin{itemize}
        \item $x = r\cos\theta$
        \item $y = r\sin\theta$
    \end{itemize} gives the same point in Cartesian coordinates.
\end{thm}

\begin{exmp}
    $(2, 2)$ in Cartesian coordinates is $(\sqrt{4 + 4}, \arctan(\frac{2}{2})) = (2\sqrt{2}, \frac{\pi}{4})$ in polar coordinates.
\end{exmp}

Since $(r, \theta + 2\pi{n})$ and $(-r, \theta + 2\pi{n})$ represent the same point in polar coordinates for all $n \in \Z$, it is common to choose ranges for $r$ and $\theta$ to make the representation of any given point unique. One such scheme is to restrict $\theta$ to $\left[0, 2\pi\right)$, and $r > 0$ except for the unique origin, $(0, 0)$.

While the two-dimensional Cartesian coordinate system is extended to three dimensions by adding the $z$-axis, there are two different ways to extend the polar coordinate system to three dimensions: cylindrical and spherical coordinates.

In cylindrical coordinates, polar coordinates are augmented with a $z$-axis, so points are represented as $(r, \theta, z)$. The radius, $r$, is still measured within the $xy$ plane.

Let $(x, y, z)$ be a point in the Cartesian coordinate system, and $(r, \theta, z)$ be the same point in cylindrical coordinates. Then all the same relationships are true as for polar coordinates, with the addition that $z = z$.

\begin{exmp}
    $(2, 2, 5)$ in Cartesian coordinates is $(\sqrt{4 + 4}, \arctan(\frac{2}{2}), 5) = (2\sqrt{2}, \frac{\pi}{4}, 5)$ in cylindrical coordinates.
\end{exmp}

In spherical coordinates, polar coordinates are augmented with a zenith angle $\phi$, measured from the zenith ($z$-axis) down to the point. The radius is measured in three dimensions rather than just in the $xy$-plane, so it is typically represented using the symbol $\rho$ rather than $r$ to differentiate it from the polar radius.

\begin{thm}
    Let $(x, y, z)$ be a point in Cartesian coordinates. Then $(\rho, \theta, \phi)$, where
    \begin{itemize}
        \item $\rho = \sqrt{x^2 + y^2 + z^2}$
        \item $\theta = \arctan\frac{y}{x}$
        \item $\phi = \arccos\frac{z}{\rho}$
    \end{itemize} gives the same point in spherical coordinates.
\end{thm}

\begin{thm}
    Let $(\rho, \theta, \phi)$ be a point in spherical coordinates. Then $(x, y, z)$, where
    \begin{itemize}
        \item $r = \rho\sin\phi$
        \item $x = r\cos\theta$
        \item $y = r\sin\theta$
        \item $z = \rho\cos\phi$
    \end{itemize} gives the same point in Cartesian coordinates.
\end{thm}

\begin{exmp}
    If $(\sqrt{6}, -\sqrt{2}, -2\sqrt{2})$ is a point in Cartesian coordinates, then $\rho$ is equal to $\sqrt{6 + 2 + 8} = 4$, $\theta = \arctan\frac{-\sqrt{2}}{\sqrt{6}} = \frac{11\pi}{6}$, and $\phi = \arccos\frac{-2\sqrt{2}}{4} = \frac{3\pi}{4}$.
\end{exmp}

\section{Surfaces}

\begin{defn}
    A \emph{level set} of a real-valued function for a value $c \in \R$ is the set of all points where the value of the function is $c$. For a function $f$ of $n$ real variables, the level set is \[L_c(f) = \left\{\left(x_1, x_2, \ldots, x_n\right) \compbar f(x_1, x_2, \ldots, x_n) = c\right\}.\]
\end{defn}

\begin{rmk}
    When $n = 2$, the level sets are called \emph{level curves}. These as essentially the same as contour lines on a topographical map. When $n=3$, the level sets are called \emph{level surfaces}.
\end{rmk}

\begin{defn}
    A \emph{cylinder} is a surface that consists of parallel lines that intersect a given plane curve. For a cylinder in the more typical sense, the plane curve is a circle.
\end{defn}

\begin{defn}
    \emph{Quadric surfaces} are higher-dimensional generalizations of conic sections --- ellipses, parabolas, and hyperbolas. In three dimensions with coordinate variables $x, y, z$, quadric surfaces are given by $Ax^2 + By^2 + Cz^2 + Dxy + Exz + Fyz + Gx + Hy + Iz + J = 0$.
\end{defn}

\begin{defn}
    An \emph{ellipsoid} with width ($x$) $a$, height ($y$) $b$, and depth ($z$) $c$ is \[\frac{x^2}{a^2} + \frac{y^2}{b^2} + \frac{z^2}{c^2} = 1.\] This can be decomposed into three orthogonal ellipses by individually setting each coordinate variable to zero.
\end{defn}

\begin{defn}
    A \emph{cone} oriented around the $z$-axis is given by \[\frac{x^2}{a^2} + \frac{y^2}{b^2} = \frac{z^2}{c^2}.\] This can be decomposed into ellipses along the $z$-axis, and pairs of lines on the $x$ and $y$ axes.
\end{defn}

\begin{defn}
    A \emph{paraboloid} oriented around the $z$-axis is given by \[\frac{x^2}{a^2} + \frac{y^2}{b^2} = \frac{z}{c}.\] This can be decomposed into ellipses around the $z$-axis, and parabolas along the $x$ and $y$ axes.
\end{defn}

\end{document}
